---
title: "ETLotto_ETLearning_analysis"
author: Emily Wertheimer
---

## LEARNING ANALYSIS
```{r}
rm(list = ls())
```

# load libraries
```{r}
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(dplyr)
```

# load data (IMPORT BY HAND)
```{r}
# setwd("C:/Users/ekw28/OneDrive - Yale University/Desktop/ETLotto_ETLearning/data")

# data_raw <- read.csv("raw_data.csv", header = TRUE)
# data_unrewarded <- read.csv("cleaned_data_unrewarded.csv", header = TRUE)
# data_summary <- read.csv("all_subs_summary_data.csv", header = TRUE)

trial_summary <- learning_trial_summary_2
learning_data_raw <- learning_data_raw_csv
learning_unrewarded <- learning_data_unrewarded2
# lotto_data_raw <- lotto_data_raw_csv

```

# create/fill matrices for ratings in unrewarded acq trials (acq cs+, acq cs-)
```{r}
thresholdAcq <- 14 

# create acq ratings matrices
ratings_acq_csp <- data.frame(matrix(nrow=100,ncol=thresholdAcq))
ratings_acq_csm <- data.frame(matrix(nrow=100,ncol=thresholdAcq))

for (i in unique(learning_unrewarded$id)){
  subjData <- filter(learning_unrewarded, learning_unrewarded$id == i)
  subjData$rating[subjData$rating==0] <- NA
  acqCSP <- filter(subjData, acq_rev == 1, csp_csm == 1)
  acqCSM <- filter(subjData,acq_rev == 1, csp_csm == 0)
  
  for (j in 1:ncol(ratings_acq_csp)) {
    ratings_acq_csp[i,j] <- acqCSP$rating[j]
  }

  for (j in 1:ncol(ratings_acq_csm)){
    ratings_acq_csm[i,j] <- acqCSM$rating[j]
  }
}
```

# create/fill matrices for ratings in unrewarded rev trials (rev cs+, rev cs-)
```{r}
thresholdRev <- 7 # change this number to change the set 

# create rev ratings matrices
ratings_rev_csp <- data.frame(matrix(nrow=100,ncol=thresholdRev))
ratings_rev_csm <- data.frame(matrix(nrow=100,ncol=thresholdRev))

# loop through all subs, retrieve current sub's data, filter by conditions, get ratings, put into appropriate ratings_rev
for (i in unique(learning_unrewarded$id)){
  subjData <- filter(learning_unrewarded, id == i)
  subjData$rating[subjData$rating==0] <- NA
  revCSP <- filter(subjData, acq_rev == 0, csp_csm == 1)
  revCSM <- filter(subjData,acq_rev == 0, csp_csm == 0)
  
  for (j in 1:ncol(ratings_rev_csp)) {
    ratings_rev_csp[i,j] <- revCSP$rating[j]
  }

  for (j in 1:ncol(ratings_rev_csm)){
    ratings_rev_csm[i,j] <- revCSM$rating[j]
  }
}

```

# write matrices
```{r}
write.csv(ratings_acq_csp, file="ratings_acq_csp.csv")
write.csv(ratings_acq_csm, file="ratings_acq_csm.csv")
write.csv(ratings_rev_csp, file="ratings_rev_csp.csv")
write.csv(ratings_rev_csm, file="ratings_rev_csm.csv")

```

## MODELING

RESCORLA-WAGNER MODEL:   ∆V = αβ(λ − ΣV)
  - learning occurs when there is an error and prediction
      Change in associative strength = learning rate(|unconditioned stimulus|-sum of associative strengths for all CS's)
      Change in associative strength= learning rate(prediction error)
  - quantifies change in associative strength of a stim after any given trial
  - prediction error is weighted by alpha
      alpha is a constant between 0-1; higher = faster learning and vice versa

deltaV = change in associative strength
alpha = learning rate (0-1; higher alpha = higher learning rate)
lambda = absolute value of unconditioned stim
vTotal = sum of all strengths for CSs
pe = lambda - vTotal # prediction error 

# SUB 1
```{r}
# reformat data 
nTrials <- as.numeric(70)
learning_rf_sub1 <- data.frame(matrix(nrow = nTrials, ncol = 7)) # learning reformatted for HJ's scripts
cols <- c("Stage", "Number", "ID", "Rating", "Stimulus", "Reinforce", "Phase")
colnames(learning_rf_sub1) <- cols

# col 1: Stage (acq/rev)
for (i in 1:nTrials) { # for sub1
  # identify reversal index
  nTrials <- 70
  first_reward_trial <- match(1, learning_data_raw$rectValue == 6) # trial num of first reward
  first_reward_ori <- sub_oris[first_reward_trial] # stim ori @ first reward
  rev_trial <- match(1,sub_oris != first_reward_ori & sub_rewards != 0)
                               
  # label Stage col as acq/rev   1: rev index -1 = acq; rev index:end = rev
  vecAcq <- vector(mode = "character", length = nTrials)
  vecAcq[1:rev_trial-1] <- "Acq"
  vecAcq[rev_trial:length(vecAcq)] <- "Rev"
  
  # put vecAcq into proper place in learning_rf$Stage
  learning_rf_sub1$Stage <- vecAcq
  
}

# col 2: Number(1-100)
learning_rf_sub1$Number <- learning_data_raw$id[1:70] # id (1-100)

# col 3: ID (10-120)
learning_rf_sub1$ID <- learning_data_raw$subjNum[1:70] # subj num (10-120)

# col 4: Rating
learning_rf_sub1$Rating <- learning_data_raw$rating[1:70]

# col 5: Stimulus (A = CS+/B = CS-)
sub1Data <- learning_data_raw[1:70,]
for (i in sub1Data$id[1:70]) {
  currentID <- sub1Data$id[i]
  sub_indices <- sub1Data$id == currentID
  sub_oris <- sub1Data$rectOri[sub_indices]
  sub_rewards <- sub1Data$rectValue[sub_indices]
  first_reward_trial <- match(1, sub_rewards == 6) 
  first_reward_ori <- sub1Data$rectOri[first_reward_trial] 

  vecStim <- ifelse(sub_oris == first_reward_ori, "A", "B")
  
  learning_rf_sub1$Stimulus[sub_indices] <- vecStim
}

# col 6: Reinforce (TRUE/FALSE)
reward1 <- learning_data_raw$rectValue[1:70]
vecReinforce1 <- ifelse(reward1 > 0, "TRUE", "FALSE")
learning_rf_sub1$Reinforce <- vecReinforce1

# label col 7 as acq1, acq2, rev1, rev2 --> HARD CODE FOR SUB 1
learning_rf_sub1$TrialNum <- learning_data_raw$trialNum[1:nTrials]
learning_rf_sub1$Phase[1:18] <- "Acq1" # 1:18 = first half of acq
learning_rf_sub1$Phase[19:37] <- "Acq2" # 19-37 = second half acq
learning_rf_sub1$Phase[38:55] <- "Rev1" # 38-54 = first half rev
learning_rf_sub1$Phase[56:70] <- "Rev2" # 55-70 = second half rev

# write sub1 reformatted df (use in Hyojung's Matlab functions for fitting)
write.csv(learning_rf_sub1, file="learning_rf_sub1.csv")

```


# for all subs
```{r}
# create df: 7000x7
  #cols: stage (Acq/Rev), Number (subjNum), ID (subjID), Rating, Stimulus (A = CS+, B = CS-), Reinforce (TRUE/FALSE),     # Phase (Acq1, Acq2, Rev1, Rev2)
learning_rf <- data.frame(matrix(nrow = nrow(learning_data_raw), ncol = 7)) # learning reformatted for HJ's scripts
cols <- c("Stage", "Number", "ID", "SP", "Stimulus", "Reinforce", "Phase")
colnames(learning_rf) <- cols
```

```{r}
# col 3: ID 
learning_rf$ID <- learning_data_raw$id # id (1-110) 

# col 2: trial num
learning_rf$Number <- learning_data_raw$trialNum # id (1-100)

# col 6: Reinforce (TRUE/FALSE)
reward <- learning_data_raw$rectValue
vecReinforce <- ifelse(reward > 0, "TRUE", "FALSE")
learning_rf$Reinforce <- vecReinforce

```

```{r}
# col 1: Stage (Acq/Rev)
for (i in unique(learning_rf$ID)) { # for e/ sub
  currentID <- learning_rf$ID[i] # set curent id
  sub_indices <- learning_rf$ID == currentID # get sub indices
  sub_oris <- learning_data_raw$rectOri[sub_indices] # get stim oris @ sub indices
  sub_rewards <- learning_data_raw$rectValue[sub_indices] # get rewards @ sub indices 
  
  # identify reversal index
  nTrials <- 70
  first_reward_trial <- match(1, sub_rewards == 6) # trial num of first reward
  first_reward_ori <- sub_oris[first_reward_trial] # stim ori @ first reward
  rev_trial <- match(1,sub_oris != first_reward_ori & sub_rewards != 0)
                               
  # label Stage col as acq/rev   1: rev index -1 = acq; rev index:end = rev
  
  learning_rf$Stage[sub_indices[1:rev_trial - 1]] <- "Acq"
  learning_rf$Stage[sub_indices[rev_trial:length(sub_indices)]] <- "Rev"
  
  
  
}

```



```{r}
# col 4: Rating (called SP)
learning_rf$SP <- learning_data_raw$rating
```

```{r}
# col 5: Stimulus (A = CS+/B = CS-)
for (i in learning_data_raw$id) {
  currentID <- learning_data_raw$id[i]
  sub_indices <- learning_data_raw$id == currentID
  sub_oris <- learning_data_raw$rectOri[sub_indices]
  sub_rewards <- learning_data_raw$rectValue[sub_indices]
  first_reward_trial <- match(1, sub_rewards == 6) 
  first_reward_ori <- learning_data_raw$rectOri[first_reward_trial] 

  vecStim <- ifelse(sub_oris == first_reward_ori, "A", "B")
  
  learning_rf$Stimulus[sub_indices] <- vecStim
}
```


```{r}
# col 7: Phase(Acq1, Acq2/ Rev1, Rev2)
for (i in unique(learning_rf$ID){ # for e/ sub
  currentID <- learning_rf$ID[i]
  sub_indices <- learning_rf$ID == currentID
  nAcq <- sum(learning_rf$Stage[sub_indices] == 'Acq')
  nRev <- sum(learning_rf$Stage[sub_indices] == 'Rev')
  learning_rf$Phase[1:nAcq/2-1]
  
  # find num of acq trials
  # find num rev trials 
  # num acq trials/2
  #1:nAcq/2-1 <- Acq1
  #nAcq:end <- Acq2
  #num rev trials/2
  #1:nRev <- Rev1
  #nRev:end <- Rev2
  
}


# learning_rf$TrialNum <- learning_data_raw$trialNum

# for (i in unique(learning_data_raw$id)){
#   currentID <- learning_data_raw$id[i] 
#   sub_indices <- learning_data_raw$id == currentID
#   
#   stageVecAcq <- vector(mode = "character", length = length(learning_rf$Stage[sub_indices] == "Acq"))
#   stageVecRev <- vector(mode = "character", length = length(learning_rf$Stage[sub_indices] == "Rev"))
# 
#   subAcq <- learning_rf$Stage[sub_indices] == "Acq"
#   subRev <- learning_rf$Stage[sub_indices] == "Rev"
#   
#   nAcq <- length(subAcq)
#   nRev <- subRev <-length(subRev)
# 
#   stageVecAcq <- ifelse(learning_rf$TrialNum[sub_indices] < nAcq/2, "Acq1", "Acq2")
#   stageVecRev <- ifelse(learning_rf$TrialNum[sub_indices] < nRev/2, "Rev1", "Rev2")
#   
#   stageVec <- c(stageVecAcq,stageVecRev)
#   learning_rf$Stage[sub_indices] <- stageVec
#   
}
```

